---
title: "Telco customer churn dataset"
output:
  html_document:
    code_folding: hide
  pdf_document: default
editor_options:
  markdown:
    wrap: sentence
---

##### Lapse risk modelling in insurance: a Bayesian mixture approach

##### Authors: Viviana G R Lobo, Thais C O Fonseca and Mariane B Alves

##### Annals of Actuarial Science

##### Manuscript ID AAS-2022-0045

This document presents a discussion descriptive statistical analysis for the Telco customer dataset shown in Section 3.3.

### Load packages and the dataset

```{r}
source("R/up(telco).R")
```

### About Telco customer dataset

The Telco customer churn data is available in IBM Business Analytics Community Connect XXXX and contains information about a Telco company that provides home phone and internet services to 7,043 customers in California in the United States.
The dataset contains 18 variables about the profile of the customers and a variable indicating who have left or stayed the service and presents an approximate 73.4% censorship rate.

```{r ,layout="l-body-outset"}
lapse1=df %>%
  group_by(lapse=Churn) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))
rmarkdown::paged_table(lapse1)
```

The response variable **tenure** represents the number of months the customer has stayed with the company with average of 32 months.
Lifetimes equal zero were removed resulting in a dataset of 7,032 customers.
Demographics characteristics are included for each customer, as well as customer account information and service information.
Description of the variables as follows:

***Demographic Information***

-   [Gender]{.underline}: Whether the client is a female or a male (Female, Male).

-   [SeniorCitizen]{.underline}: Whether the client is a senior citizen or not ( 0, 1).

-   [Partner]{.underline}: Whether the client has a partner or not (Yes, No).

-   [Dependents]{.underline}: Whether the client has dependents or not (Yes, No).

***Services Information***

-   [PhoneService]{.underline}: Whether the client has a phone service or not (Yes, No).

-   [MultipleLines]{.underline}: Whether the client has multiple lines or not (No phone service, No, Yes).

-   [InternetServices]{.underline}: Whether the client is subscribed to Internet service with the company (DSL, Fiber optic, No)

-   [OnlineSecurity]{.underline}: Whether the client has online security or not (No internet service, No, Yes).

-   [OnlineBackup]{.underline}: Whether the client has online backup or not (No internet service, No, Yes).

-   [DeviceProtection]{.underline}: Whether the client has device protection or not (No internet service, No, Yes).

-   [TechSupport]{.underline}: Whether the client has tech support or not (No internet service, No, Yes).

-   [StreamingTV]{.underline}: Whether the client has streaming TV or not (No internet service, No, Yes).

-   [StreamingMovies]{.underline}: Whether the client has streaming movies or not (No internet service, No, Yes).

***Customer Account Information***

-   [tenure]{.underline}: Number of months the customer has stayed with the company (Multiple different numeric values).

-   [Contract]{.underline}: Indicates the customer's current contract type (Month-to-Month, One year, Two year).

-   [PaperlessBillin]{.underline}g: Whether the client has paperless billing or not (Yes, No).

-   [PaymentMethod]{.underline}: The customer's payment method (Electronic check, Mailed check, Bank transfer (automatic), Credit Card (automatic)).

-   [MontlyCharges]{.underline}: The amount charged to the customer monthly (Multiple different numeric values).
    ´

-   [TotalCharges]{.underline}: The total amount charged to the customer (Multiple different numeric values).

```{r, layout="l-body-outset"}
rmarkdown::paged_table(df)
```

### Descriptive Analysis via Kaplan-Meier

The panels illustrate the pattern of the lifetime of the customers.
Although a large admission of new customers is observed for the first months of contract, Third panel indicates there is a high lapse rate at the beginning, i.e., there is a large portion of customers that will leave after just a few months of service.
In the other months, the lapse risk rate remains around 0.005.
Note that after 60 months of contract, there is a steep increase in the lapse rate.
We expect that our proposed model is able to capture these sudden changes of pattern in the hazard function over time.

We present a descriptive analysis based on empirical marginal survival curves with the Kaplan-Meier estimator and evaluate the hazard curve for some covariates including the categories demographics, service information, and customer account information.
As we can be seen, categories for each variable present particular behaviour and could be useful to understand the time to churn.
Although there is a large set of covariates available to explain the lapse risk, a group of them is not informative for the survival time under study.

```{r}
dados <-  df
#Kaplan-Meier fit (marginal)
fit.km1 <- survfit2(Surv(tenure, status) ~ gender, dados)
fit.km2 <- survfit2(Surv(tenure, status) ~ PaymentMethod, dados)
fit.km3 <- survfit2(Surv(tenure, status) ~ Contract, dados)
fit.km4 <- survfit2(Surv(tenure, status) ~ SeniorCitizen, dados)
fit.km5 <- survfit2(Surv(tenure, status) ~ InternetService, dados)
fit.km6 <- survfit2(Surv(tenure, status) ~ StreamingTV, dados)
fit.km8 <- survfit2(Surv(tenure, status) ~ Dependents, dados)
fit.km9 <- survfit2(Surv(tenure, status) ~ PaperlessBilling, dados)
fit.km10<- survfit2(Surv(tenure, status) ~ DeviceProtection, dados)
fit.km11<- survfit2(Surv(tenure, status) ~ PhoneService, dados)
fit.km12<- survfit2(Surv(tenure, status) ~ TechSupport, dados)

extrai_km <- function(fit_km) {
    ret <- tidy_survfit(fit_km, type = "surv") %>%
    select(.eval_time = time, .pred_survival = estimate, id = strata) %>%
      group_by(id) %>%
      mutate(haz = tx.emp(.eval_time, .pred_survival)) %>%  
      tidyr::nest(.pred = c(.eval_time, .pred_survival))
      return(ret)
}

kms <- list(Gender=fit.km1, PaymentMethod=fit.km2, Contract=fit.km3, 
            SeniorCitizen= fit.km4, InternetService= fit.km5, 
            StreamingTV= fit.km6,Dependents=fit.km8, PaperlessBilling= fit.km9, DeviceProtection=fit.km10, PhoneService= fit.km11, TechSupport=fit.km12) 
resultado= map(kms, extrai_km)
```

### Gender

The lapse rate for the two categories of \texttt{Gender} in the data is nearly the same.
In this case, we do not expected that Gender is significant to explain the lapse risk.

```{r}
preds <-bind_rows(Gender=resultado$Gender, .id = "modelo") %>%
  group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)

a1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

a2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(a1) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(a2) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

**PaymentMethod**

For the payment method see that the behaviour of survival curves and churn rates for Credit Card (automatic), Bank Transfer (automatic) and Mailed Check are quite similar.
On the other hand, when we consider the instantaneous lapse curve, notice that customers that preferable for an {Eletronic check} as a payment method are the majority to leave the company.
In this case, we consider jointly these classes in a new category called \`\`others''.

```{r}
preds <-   
  bind_rows(PaymentMethod=resultado$PaymentMethod, .id = "modelo") %>%
  group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)

b1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

b2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(b1)  %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(b2) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

**Contract**

For the customer account information, see that customers with Month-to-Month contracts have higher lapse rates compared to clients with One-year and Two-year contracts.
See that the hazard curve are very similar for the contract that has payment more than one year.
In this way, we consider to aggregate in a unique group called {\`\`One-year +''}.

```{r}
preds <-   
 bind_rows(Contract=resultado$Contract, .id = "modelo") %>%
  group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)

c1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

c2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(c1) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(c2) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

### SeniorCitizen

See that no Senior citizen has lower propensity to churn when compared with senior citizen.
Although this covariate apparently significant to explain the rate of lapsation over time, the SeniorCitizen covariate does not show significant in the modelling.

```{r}
preds <-   
 bind_rows(SeniorCitizen=resultado$SeniorCitizen, .id = "modelo") %>%
 group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)

d1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

d2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(d1) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(d2) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

### InternetService

The behaviour of customers that do not have internet service or DSL service present similar behaviour in terms of lapse.
See that customer that have Fiber optic service has proposenty to churn earlier compared each others.
As the hazard curve are quite similar for No and DSL categorie, we consider to aggregate in an unique category called "others".

```{r}
preds <-   
bind_rows(InternetService=resultado$InternetService, .id = "modelo") %>%
  group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)

e1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

e2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

```

```{r, echo=FALSE}
ggplotly(e1) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(e2) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

### StreamingTV

```{r}
preds <-   
 bind_rows(StreamingTV=resultado$StreamingTV, .id = "modelo") %>%
  group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)


f1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

f2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(f1) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(f2) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

### Dependents

This covariate apparently does not enough to explain the time to churn.
See that the instantaneous lapse risk are quite similar for Yes and No categories.

```{r, echo=FALSE}
preds <-   
bind_rows(Dependents=resultado$Dependents, .id = "modelo") %>%
  group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)


g1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

g2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(g1) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(g2) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

### PaperlessBilling

This covariate apparently does not enough to explain the time to churn.
See that the instantaneous lapse risk are quite similar for Yes and No contegories.

```{r}
preds <-   
bind_rows(PaperlessBilling=resultado$PaperlessBilling, .id = "modelo") %>%
  group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)



h1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

h2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(h1) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(h2) %>% layout(legend=list(orientation="h",x=0.2, y=-0.2)) 
```

### DeviceProtection

This covariate presents collinearity with variable **InternetService**.
See the dependence in the cateogory \`\`No InternetService''.

```{r}
preds <-   
 bind_rows(DeviceProtection=resultado$DeviceProtection, .id = "modelo") %>%
 group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)


i1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

i2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(i1) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(i2) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

### PhoneService

Similar behaviour has already been mentioned for the variable **Gender**.

```{r}
preds <-   
 bind_rows(PhoneService=resultado$PhoneService, .id = "modelo") %>%
 group_by(modelo) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred)


j1<- ggplot(aes(x = .eval_time, y = .pred_survival, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("survival probabilities")+
  geom_step(linewidth=1) + 
   theme_bw() +
  scale_y_continuous(limits=c(0.2,1), breaks = seq(0.2,1,by=.2), labels = point) +
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)

j2<-ggplot(aes(x = .eval_time, y = haz, col=id), data = preds) +
  xlab("survival time (in months)") + ylab("hazard rate")+
  theme_bw() +
  geom_line(linewidth=1) + 
   scale_y_continuous(limits=c(0,0.06), breaks = seq(0,0.06,by=.02), labels = point)+
  scale_x_continuous(limits=c(0,75), breaks = seq(0,75,by=20), labels = grad)
```

```{r, echo=FALSE}
ggplotly(j1) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

```{r, echo=FALSE}
ggplotly(j2) %>% layout(legend=list(orientation="h",x=0, y=-0.2)) 
```

The lapse rate for the two categories of **Gender** in the data are nearly the same.
Similar behaviour is observed for some customer service information such as **PhoneService**, **MutipleLines** and **OnlineSecurity**.
Also, there is multicollinearity for groups of variables, in particular **OnlineSecurity**, **TechSupport**, **OnlineBackup** and **DeviceProtection** are all dependent on the **OnlineService** variable.
Also, the covariates **MonthlyCharges** and **TotalCharges** present high collinearity.

```{r}
sp1= ggplot(aes(x=MonthlyCharges, y=TotalCharges), data=df) +
  theme_bw() +
  geom_point()
```

```{r}
ggplotly(sp1)
```
