---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LapseRiskASS

##### Lapse risk modelling in insurance: a Bayesian mixture approach
##### Authors: Viviana G R Lobo, Thais C O Fonseca and Mariane B Alves


Bayesian Survival models via mixture of Log-Normal distribution extends the well-known survival models accommodating different behaviour over time and considers higher censored survival times. The proposal combines mixture distributions Fruhwirth-Schnatter(2006) <https://doi.org/10.1007/s11336-009-9121-4>, and data augmentation techniques Tanner and Wong (1987) <https://doi/abs/10.1080/01621459.1987.10478458>.

More details about the lnmixsurv package see:
<https://github.com/vivianalobo/lnmixsurv.git>

## Requirements

The codes use the Bayesian mixture survival approach implemented in the lnmixsurv package. 

```{r}
require("lnmixsurv")
```

## Example

We present the Telco customer analysis. This is a basic example which shows you how to adress the dataset and run descriptive analysis and inference procedure

```{r}
#### Telco customer churn dataset 
### loading data and others
source("R/up(telco).R")
```

For see a detailed descriptive analysis ....


To run the inference procedure see ....



